---
title: "Regression Example"
author: "Ken Oung Yong Quan"
date: "1 August 2016"
output: html_document
---

We'll be using the mtcars dataset for this example.

It's available in R's built-in base package. Retrieve it using the `data(mtcars)`.

```{r}
# Importing the data
data(mtcars)

library(data.table)
mtcars <- as.data.table(mtcars)

# Find out more about this dataset
?mtcars
```

Using a multiple linear regression model, we want to investigate the fuel efficiency of cars by looking at their mpg (miles per gallon) rating.

```{r}
# Fit the model
fit <- lm(mpg ~ ., data = mtcars)
summary(fit)
```

By including all the variables, we get a decent model with an adjusted R-squared of 0.8066. However, we note that the p-values are all rather high. This indicates that the model can be improved by removing some of these regressors. 

```{r}
# Removing the regressor with the highest p-value
fit_no_cyl <- lm(mpg ~ . - cyl, data = mtcars)
summary(fit_no_cyl)
```

Note the improvement in adjusted R-squared. We repeat this until all our regressors have p-values below 0.05.

```{r}
fit_improved <- lm(mpg ~ wt + qsec + am)
summary(fit_improved)
```

This process can actually be automated using the stepAIC function in the MASS package.

```{r}
library(MASS)
stepAIC(fit)
```

However, there are actually better models that you can obtain by applying some simple transformations. For example, note that the relationship between mpg and hp isn't exactly linear. We can obtain a more linear relationship by applying a log transform to the hp column.

```{r}
# Before transform
plot(mtcars[, mpg], mtcars[, hp])

# After transform 
plot(mtcars[, mpg], log(mtcars[, hp]))
```

Try this new adjusted hp in this regression model.

```{r}
fit_transformed <- lm(mpg ~ . - hp + log(hp), data=mtcars)
stepAIC(fit_transformed)

fit_final <- lm(mpg ~ wt + am + log(hp), data=mtcars)
summary(fit_final)
```

```{r}
# Regression Assumptions
plot(fit_final)

# Check for multi-collinearity
# Anything below 10 is okay
vif(fit_final)
```